Near and Mid-infrared properties of known $z\geq5$ Quasars

### 2019-06-11 tex/VHzQ.tex
color_redshift/SpecType re-jig
things_still_todo

@@

\begin{table}
  \centering
  \begin{tabular}{l r r r r}
    \hline \hline
    \multirow{ 2}{*}{$z \geq$} & \multicolumn{2}{c}{unWISE} &  \multicolumn{2}{c}{ALLWISE}   \\ 
                                             & \multicolumn{1}{c}{W1} & \multicolumn{1}{c}{W2} & \multicolumn{1}{c}{W3} & \multicolumn{1}{c}{W4}  \\ 
    \hline
    5.00   &   362 ( 78.2\%)   &   308 ( 66.5\%)  &    51 ( 11.0\%)  &    10 (  2.2\%) \\ 
    5.70   &   186 ( 69.4\%)   &   151 ( 56.3\%)  &    15 (  5.6\%)  &     2 (  0.7\%) \\ 
    6.00   &   109 ( 63.7\%)   &    91 ( 53.2\%)  &     8 (  4.7\%)  &     2 (  1.2\%) \\ 
    6.19   &    63 ( 72.4\%)   &    47 ( 54.0\%)  &     3 (  3.4\%)  &     2 (  2.3\%) \\ 
    6.50   &    34 ( 85.0\%)   &    26 ( 65.0\%)  &     1 (  2.5\%)  &     2 (  5.0\%) \\ 
    6.78   &    12 ( 85.7\%)   &     9 ( 64.3\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    7.00   &     4 (100.0\%)   &     4 (100.0\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    7.50   &     1 (100.0\%)   &     1 (100.0\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    \hline \hline
  \end{tabular}
  \caption{Detection rates in the mid-infrared bands from WISE as a function of 
    redshift. A signal-to-noise cut of {\tt w$x$snr}$>3.0$ has been used for each band.}
  \label{tab:WFCAM_vs_VISTA}
\end{table}

@@

\citet{Blain2013}

@@

\begin{figure}
    \centering
    \includegraphics[width=8.6cm,  clip, trim=6mm 4mm 6mm 10mm]
    {../color_color/W1W2_plots/W1W2W3_hexplots_20190610v2.png} 
  \vspace{-14pt}
    \caption{The (W2-W3) vs. (W1-W2) colour – colour diagram showing the WISE colours 
      for the 283 VH$z$Q that have reported w3mpro values (blue to red coloured points).
      400,000 quasars at redshift $z\lesssim3.0$ from the DR14Q are also shown
      (grey colour-scale). }
    \label{fig:W1W2W3}
\end{figure}

\subsection{MIR Colours}

%% 
%%    (W1 - W2)    vs.  (W2 - W3) 
%% Call W1 ~ 3-4um annd W2   4-5um and   W3 10-16um
Figure~\ref{fig:W1W2W3} shows the (W1-W2) vs. (W2-W3) colour – colour
space diagram showing the WISE colours for the 283 VH$z$Q that have
reported w3mpro values (blue to red coloured points).  400,000 quasars
at redshift $z\lesssim3.0$ from the DR14Q are also shown (grey
colour-scale). 
W1 samples $\approx$0.6-4$\mu$m for the $z=0-3$ quasars, 
W2 samples $\approx$1.25-5$\mu$m for the $z=0-3$ quasars, 
W3 samples $\approx$2.-16$\mu$m for the $z=0-3$ quasars, 
1

%%
The VH$z$Q with the bluest colour in (W1 - W2) is 
DELS J1048-0109 with (W1-W2) = 0.154 
(see Figure~\ref{fig:DELS_J1048}).

%\subsubsection{Individual VH$z$Qs with Interesting MIR properties}
Very High-$z$ Quasars Detected in WISE W3 and W4.

\begin{figure*}
    \centering
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/UHS_J0439+1634.png} 
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J0100+2802.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J1443+3623.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J1623+4705.png}
    \caption{ (a) UHS J0439+1634 
                 (b) SDSS J0100+2802 
               (c) SDSS J1443+3623 
 (d) SDSS J1623+4705
}
    \label{fig:foobar}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=8.5cm]
  {/cos_pc19a_npr/programs/quasars/highest_z/light_curves/MIR_LCs/NEOWISER_LC_histogramlog_20180827.png}
  \vspace{-16pt}
  \caption[]
  {Histogram showing the number of NEOWISE-R epochs and detections there are for each 
    VH$z$Q.} 
  \label{fig:MIR_LC_epochs}
\end{figure}

\subsection{Variability}
VH$z$Qs, if accreting at, or above the Eddington Limit, might well
have have large values of changing mass accretion rate, $\ddot{m}_{\rm
accr}$. A consequence of this would be that these quasars exhibit signs
of variability, most likely showing up in their UV/optical rest-frame
spectra. We look for evidence of this variability signature in the NIR
and MIR light-curves of the VH$z$Qs. As a guide, \civ enters the
$Y$-band at redshift $z$=5.32 and exits at $z$=5.99, and enters the
$J$-band at redshift $z=6.55$ and exits at $z$=7.57. \mgii enters the
$H$-band at redshift $z=4.33$ and exits at $z=5.37$ and enters the
$K$-band at redshift $z=6.25$ and exits at $7.50$.

@@

Figure~\ref{fig:MIR_LC_epochs} gives the number of NEOWISE-R epochs and detections there are for each VH$z$Q, while 
Figure~\ref{fig:MIR_LC_3egs} presents three examples of the MIR lightcurves and
associated colour changes. Here we show J0100+2802 \citep{Wu2015}, J0224-4711 and  J1626+2751. 
{\bf NJC: What about NIR light-curves / combined light-curves}

At least 8 original observations across at least 30 days in either WSA or VSA per band, calibrated flux greater than zero over the whole time, and the error on the average flux is 8 times lower than this.

Then a requirement of SNR$>3$ per `average' epoch, where `average epoch' is then dependent on the averaging timescales (which as we see can vary).

@@

Longer time-series, better SNR and moreover, $(1+z)$. 
Just selecting things that would
%%
Doing this independently in each band...
%Variation

%% 
%% Object numbers: 54 VSA   MMTJ0215-0529,
%% objID: 162 is both VSA and WSA and  SDSSJ0959+0227
%%
objiD 164 is SDSSJ1000+0234, faint; very faint. 
SHELLQsJ0220-0432: (qso59) show this too. 
CFHQSJ0216-0455

@@

VHS, Area 1: 150$<$RA/deg$<200$ and $-40<$Decl./deg$<-20$:
$Y$=21.1 mag, $J$=20.6 mag, $H$=20.3 mag, $K_{s}$=19.8 mag (all AB). 
Area 2: 10$<$RA/deg$<$60, $-60<$Dec$<-40$:
$J$=21.2 mag, $H$=20.6 mag, $K_{s}$=20.2 mag.

VIKING is more uniform, so I calculated the averages over the whole area:
$Z$=22.7 mag, $Y$=21.9 mag, $J$=21.4 mag, $H$=21.2 mag, $Ks$=21.1 mag.

\begin{figure}
  \centering
  \includegraphics[width=8.5cm]{../data/WSA_VSA/VHS_J_abMagLim.png}
  \caption{The calculated average {\tt AB MAGLIM} in the VHS.
    As one can see, VHS is not completely uniform. The are 3 areas with different filter sets, 
    and the depths changed in 2 of these areas, particularly in $J$ and $K_{s}$. 
    Y and H are almost uniform, but the coverage is much less.}
  \label{fig:1}
\end{figure}

Following \citet{Fan2001b} and \citet{McGreer2013}, we use use an
exponential decline to describe the space density of VH$z$Qs at high
redshifts, e.g.
\begin{equation}
\rho(z, {\rm M}_{1450} \propto 10^{k (z)}
\end{equation}
where $z$ is the sample redshift. 
%Fan, X., Strauss, M. A., Schneider, D. P., et al. 2001b, AJ, 121, 54

UHS J-band depth 19.6 \citep[Vega; ][]{Dye2018}

VISTA HEMISPHERE SURVEY DATA RELEASE 1
Release date (will be set by ESO)
PROPOSAL ESO No.: 179.A-2010
PRINCIPAL INVESTIGATOR: Richard McMahon
Authors: R. McMahon, M. Banerji, N. Lodieu for the VHS Collaboration

The aim of the Vista Hemisphere Survey (VHS) is to carry out a near
Infra-Red survey, which when combined with other VISTA Public Surveys
will result in coverage of the whole southern celestial hemisphere
($\sim$20,000 deg$^2$) to a depth 30 times fainter than 2MASS/DENIS in
at least two wavebands (J and K$_s$), with an exposure time of 60 seconds
per waveband to produce median 5σ point source (Vega) limits of $J =
20.2$ and $K_{S} = 18.1$. In the South Galactic Cap, $\sim$5000 deg$^{2}$ will be
imaged deeper with an exposure time of 120 seconds and also including
the H band producing median 5$\sigma$ point limits of: J = 20.6; H = 19.8; K$_{s}$ = 18.5. 
In this 5000 deg$^{2}$ region of sky deep multi-band optical (grizY)
imaging data will be provided by the Dark Energy Survey (DES). The
remainder of the high galactic latitude ($|b|>$30$^{\circ}$) sky will be imaged
in YJHK for 60sec per band to be combined with ugriz waveband
observations from the VST ATLAS survey. \\

VHS                   YJHK            99.99, 21.2,  21.2,  20.6,  20.0                (9) \\

{\tt ABmagLimits} for each survey in the database. \\

>>>>

Figure~\ref{fig:coverage} shows the areal coverage of the ULAS, UHS,
VHS and VIKING. UHS DR1 is 12,600 deg$^{2}$ ($J$-band); ULAS DR11
3,700 deg$^{2}$ VHS DR6 16,000 deg$^{2}$ and VIKING DR5 is 1,300
deg$^{2}$. The overlap between UHS DR1 and VHS DR6 is 28 deg. These
four surveys together cover 33,000 deg$^{2}$.

@@

\begin{table}
  \centering
  \begin{tabular}{l r r r r}
    \hline \hline
    \multirow{ 2}{*}{$z \geq$} & \multicolumn{2}{c}{unWISE} &  \multicolumn{2}{c}{ALLWISE}   \\ 
                                             & \multicolumn{1}{c}{W1} & \multicolumn{1}{c}{W2} & \multicolumn{1}{c}{W3} & \multicolumn{1}{c}{W4}  \\ 
    \hline
    5.00   &   362 ( 78.2\%)   &   308 ( 66.5\%)  &    51 ( 11.0\%)  &    10 (  2.2\%) \\ 
    5.70   &   186 ( 69.4\%)   &   151 ( 56.3\%)  &    15 (  5.6\%)  &     2 (  0.7\%) \\ 
    6.00   &   109 ( 63.7\%)   &    91 ( 53.2\%)  &     8 (  4.7\%)  &     2 (  1.2\%) \\ 
    6.19   &    63 ( 72.4\%)   &    47 ( 54.0\%)  &     3 (  3.4\%)  &     2 (  2.3\%) \\ 
    6.50   &    34 ( 85.0\%)   &    26 ( 65.0\%)  &     1 (  2.5\%)  &     2 (  5.0\%) \\ 
    6.78   &    12 ( 85.7\%)   &     9 ( 64.3\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    7.00   &     4 (100.0\%)   &     4 (100.0\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    7.50   &     1 (100.0\%)   &     1 (100.0\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    \hline \hline
  \end{tabular}
  \caption{Detection rates in the mid-infrared bands from WISE as a function of 
    redshift. A signal-to-noise cut of {\tt w$x$snr}$>3.0$ has been used for each band.}
  \label{tab:mir_detection}
\end{table}

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=8.6cm,  clip, trim=6mm 4mm 6mm 10mm]
    {../color_color/W1W2_plots/W1W2W3/W1W2W3_hexplots_20190610v2.png} 
  \vspace{-14pt}
    \caption{The (W2-W3) vs. (W1-W2) colour – colour diagram showing the WISE colours 
      for the 283 VH$z$Q that have reported w3mpro values (blue to red coloured points).
      400,000 quasars at redshift $z\lesssim3.0$ from the DR14Q are also shown
      (grey colour-scale). }
    \label{fig:W1W2W3}
\end{figure}

\subsection{MIR Colours}
%% The VH$z$Q with the bluest colour in (W1 - W2) is 
DELS J1048-0109 with (W1-W2) = 0.154. 
%(see Figure~\ref{fig:DELS_J1048}).
%%
%%    (W1 - W2)    vs.  (W2 - W3) 
%%
%%  Call W1 ~ 3-4um annd W2   4-5um and   W3 10-16um
Figure~\ref{fig:W1W2W3} shows the (W1-W2) vs. (W2-W3) colour – colour
space diagram showing the WISE colours for the 283 VH$z$Q that have
reported w3mpro values (blue to red coloured points).  400,000 quasars
at redshift $z\lesssim3.0$ from the DR14Q are also shown (grey
colour-scale). 
% W1 samples $\approx$0.6-4$\mu$m for the $z=0-3$ quasars, 
% W2 samples $\approx$1.25-5$\mu$m for the $z=0-3$ quasars, 
% W3 samples $\approx$2.-16$\mu$m for the $z=0-3$ quasars,

%%
%\subsubsection{Individual VH$z$Qs with Interesting MIR properties}

\begin{figure*}
    \centering
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/UHS_J0439+1634.png} 
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J1443+3623.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J1623+4705.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J0100+2802.png}
    \caption{ (a) UHS J0439+1634 
                 (b) SDSS J0100+2802 
               (c) SDSS J1443+3623 
 (d) SDSS J1623+4705
}
    \label{fig:VHzQ_W4s}
\end{figure*}
The are a set of VH$z$Q detected in WISE W3 and W4. 51 objects are
detected in the broad W3 filter and 10 objects are formally detected
in W4 ({\tt w4snr}$\geq$3.0). Four of the brightest W4 objects are
presented in \label{fig:VHzQ_W4s}. UHS J0439+1634
(Fig.~\ref{fig:VHzQ_W4s} (a)) with W4=7.165$\pm$0.12 was discovered by
\citet{Fan2019} and is a strongly lensed quasar at $z=$6.51. This high
luminosity is most not intrinsic, but is boosted by an intervening
redshift $z\sim$0.7 galaxy. SDSS J0100+2802 (redshift $z=6.33$) does
not have a formal W4 detection (W4 $=8.98\pm0.45$ and {\tt w4snr}=2.4)
but as reported by \citet{Wu2015}, has a bright detections in the
2MASS $J$, $H$ and $K_{\rm s}$ bands and we report it here.

\begin{figure}
  \centering
  \includegraphics[width=8.5cm]
  {/cos_pc19a_npr/programs/quasars/highest_z/light_curves/MIR_LCs/NEOWISER_LC_histogramlog_20180827.png}
  \vspace{-16pt}
  \caption[]
  {Histogram showing the number of NEOWISE-R epochs and detections there are for each 
    VH$z$Q.} 
  \label{fig:MIR_LC_epochs}
\end{figure}
\subsection{Variability}
VH$z$Qs, if accreting at, or above the Eddington Limit, might well
have have large values of changing mass accretion rate, $\ddot{m}_{\rm
accr}$. A consequence of this would be that these quasars exhibit signs
of variability, most likely showing up in their UV/optical rest-frame
spectra. We look for evidence of this variability signature in the NIR
and MIR light-curves of the VH$z$Qs. As a guide, \civ enters the
$Y$-band at redshift $z$=5.32 and exits at $z$=5.99, and enters the
$J$-band at redshift $z=6.55$ and exits at $z$=7.57. \mgii enters the
$H$-band at redshift $z=4.33$ and exits at $z=5.37$ and enters the
$K$-band at redshift $z=6.25$ and exits at $7.50$.

@@

Figure~\ref{fig:MIR_LC_epochs} gives the number of NEOWISE-R epochs and detections there are for each VH$z$Q, while 
Figure~\ref{fig:MIR_LC_3egs} presents three examples of the MIR lightcurves and
associated colour changes. Here we show J0100+2802 \citep{Wu2015}, J0224-4711 and  J1626+2751.

At least 8 original observations across at least 30 days in either WSA
or VSA per band, calibrated flux greater than zero over the whole
time, and the error on the average flux is 8 times lower than this.
%%
Then a requirement of SNR$>3$ per `average' epoch, where `average epoch' is then dependent on the averaging timescales (which as we see can vary).

@@

Longer time-series, better SNR and moreover, $(1+z)$. 
Doing this independently in each band...

%% 
%% Object numbers: 54 VSA   MMTJ0215-0529,
%% objID: 162 is both VSA and WSA and  SDSSJ0959+0227
%%

%% 
%% objiD 164 is SDSSJ1000+0234, faint; very faint. 
%% SHELLQsJ0220-0432: (qso59) show this too. 
%%

%%
%% CFHQSJ0216-0455
%%

@@

In the North, we assume the UHS J-band depth of 19.6 \citep[Vega; ][]{Dye2018}. 
In the South, one can use the report {\tt ABmagLimits} for each survey in the database, we
can calculate the depths of the various NIR VISTA surveys.
Fig.~\ref{fig:VHS_J_abMagLim} shows this calculated average {\tt AB
MAGLIM} in the VHS.  As one can see, the VHS is not completely
uniform; one area (150$<$RA/deg$<200$ and $-40<$Decl./deg$<-20$:) has
($Y$, $J$, $H$, $K_{s}$) = (21.1, 20.6, 20.3, 19.8) mag, while a
second area (10$<$RA/deg$<$60, $-60<$Dec$<-40$) has ($Y$, $J$, $H$,
$K_{s}$) = (??, 21.2, 20.6, 20.2) mag.  VIKING is more uniform, with
average depths over the whole VIKING area of ($Z$, $Y$, $J$, $H$,
$K_{s}$) = (22.7, 21.9, 21.4, 21.2, 21.1) mag.

\begin{figure}
  \centering
  \includegraphics[width=8.5cm]{../data/WSA_VSA/VHS_J_abMagLim.png}
  \caption{The calculated average {\tt AB MAGLIM} in the VHS.
    As one can see, VHS is not completely uniform. The are 3 areas with different filter sets, 
    and the depths changed in 2 of these areas, particularly in $J$ and $K_{s}$. 
    Y and H are almost uniform, but the coverage is much less.}
  \label{fig:VHS_J_abMagLim}
\end{figure}

Following \citet{Fan2001b} and \citet{McGreer2013}, we use use an
exponential decline to describe the space density of VH$z$Qs at high
redshifts, e.g.
\begin{equation}
\rho(z, {\rm M}_{1450}) \propto 10^{k (z)}
\end{equation}
where $z$ is the sample redshift. 
%Fan, X., Strauss, M. A., Schneider, D. P., et al. 2001b, AJ, 121, 54
